Lessons learned from buggy models of human history

### 2020-05-20 paper.tex
Changes and refs

\title{Lessons learned from bugs in models of human history}
\author[1]{Aaron P. Ragsdale}
\author[1]{Dominic Nelson}
\author[1,$\star$]{Simon Gravel}
\author[2,$\star$]{Jerome Kelleher}
\affil[1]{McGill University and Genome Qu\'{e}bec Innovation Centre,
McGill University, Montr\'{e}al, Qu\'{e}bec, Canada}
\affil[2]{Big Data Institute, Li Ka Shing Centre for Health Information and Discovery,
University of Oxford, Oxford, United Kingdom}
\affil[$\star$]{Joint senior authors, listed alphabetically}
\maketitle

\aprcomment{title thoughts: \\
Reducing errors in demographic model design and implementation \\
}
\sgcomment{How to avoid simulating the wrong demographic model}

\abstract{
Simulation plays a central role in population genomics studies. Recent years
have seen rapid improvements in software efficiency that make it possible to simulate
large genomic regions for many individuals sampled from large numbers of populations.
The increase in complexity of possible demographic models also provides additional ways
that we can get their implementation wrong. Here we describe two errors made in
defining population genetic models using the \msprime\ coalescent simulator that have
found their way into the published record.
We discuss how these errors have affected analyses and suggest recommendations
for software developers and users to reduce the risk of such errors.
}

@@

\section*{Case 1: A misspecified model in msprime's documentation}

@@

\section*{Case 2: Incorrect model parameters in an analysis pipeline}

@@

This simulation was performed to assess the
transferability of polygenic risk scores across human populations. In other words, it sought to
explore how human demographic history and population structure affect our ability
to predict genetic risk in diverse populations given the well-documented unequal representation
in medical genetic studies \cite{Popejoy}. The resulting publication has been influential in the discussion of
health inequalities and genomics, with over 350 citations since 2017.

The large excess in divergence under the incorrect model here exaggerated the role of demography
and genetic drift in limiting the transferability of genetic risk scores across populations.

>>>>

\title{Lessons learned from buggy models of human history}
\author[1]{Aaron P. Ragsdale}
\author[1]{Dominic Nelson}
\author[1,$\star$]{Simon Gravel}
\author[2,$\star$]{Jerome Kelleher}
\affil[1]{McGill University and Genome Qu\'{e}bec Innovation Centre,
McGill University, Montr\'{e}al, Qu\'{e}bec, Canada}
\affil[2]{Big Data Institute, Li Ka Shing Centre for Health Information and Discovery,
University of Oxford, Oxford, United Kingdom}
\affil[$\star$]{Joint senior authors, listed alphabetically}
\maketitle

\aprcomment{title thoughts: \\
Reducing errors in demographic model design and implementation \\
Lessons learned from misspecified models of human history \\
Lessons learned from buggy models of human history \\
Lessons learned from bugs in population genetics models
}

\abstract{
Simulation plays a central role in population genomics studies. Recent years
have seen rapid improvements in software efficiency that make it possible to simulate
large genomic regions for many individuals sampled from large numbers of populations.
\aprcomment{fix this sentence ->} The increase in complexity of possible demographic models also provides additional ways
that we can get their implementation wrong. Here we describe two errors made in
defining population genetic models using the \msprime\ coalescent simulator that have
found their way into the published record.
We discuss how these errors have affected analyses and suggest recommendations
for software developers and users to reduce the risk of such errors.
}

@@

\section*{A bad tutorial example}

@@

\section*{Risk prediction in human populations}

@@

This simulation was performed to assess the
transferability of polygenic risk scores across human populations. In other words, it sought to
explore how human demographic history and population structure affect our ability
to predict genetic risk in diverse populations given the well-documented unequal representation
in medical genetic studies~\citep{popejoy2016genomics}. The resulting publication has been
influential in the discussion of health inequalities and genomics, with over 350 citations since 2017.
The large excess in divergence under the incorrect model here exaggerated the role of demography
and genetic drift in limiting the transferability of genetic risk scores across populations.
